D.I.A.G.R.A.M.: Development of Image Analysis for Graph Recognition And Modeling

### 2025-07-01 doc/article/article.tex
Updated paper

In comparison to related work, D.I.A.G.R.A.M. distinguishes itself by targeting offline, image-based recognition (as opposed to stroke-based online methods) while retaining support for structural output generation. This positions it as a hybrid between traditional rule-based recognizers and modern neural approaches.

We'd like to point out a few things about Figure~\ref{fig:results}. First of all, the result shown in part (d). A non-existing node can be seen in the upper right corner of the image; this is due to the fact that the self-arrow of node $q_5$ is recognized - by the object detection network - both as an arrow and as a node, with an high trust value. This problem could have been solved by applying a Non-Maxima Suppression - but in this specific case, the highest confidence value was surprisingly the node's one. This problem - that occasionally shows up - is probably to be attributed to the low samples present in the dataset.

>>>>

In comparison to related work, D.I.A.G.R.A.M. distinguishes itself by targeting offline, image-based recognition (as opposed to stroke-based online methods) while retaining support for structural output generation. This positions it as a hybrid between traditional rule-based recognizers and modern neural approaches. \\

We'd like to point out a few things about the results shown Figure~\ref{fig:results}. First of all, the result shown in part (d). A non-existing node can be seen in the upper right corner of the image; this is due to the fact that the self-arrow of node $q_5$ is recognized - by the object detection network - both as an arrow and as a node, with an high trust value. This problem could have been solved by applying a Non-Maxima Suppression - but in this specific case, the highest confidence value was surprisingly the node's one. This problem - that occasionally shows up - is probably to be attributed to the low samples present in the dataset.

\begin{figure}[h!]
	\centering
	\vspace{0.5em}

	% First row
	\begin{tabular}{ccc}
		\includegraphics[height=5cm]{ex6.png} &
		\includegraphics[height=5cm]{chat1.png} &
		\includegraphics[height=5cm]{chat2.png} \\
		% Optional captions under each image
		(a) Our result & (b) GPT Mermaid & (c) GPT D2 \\
	\end{tabular}

	% Second row
	\begin{tabular}{ccc}
		\includegraphics[height=5cm]{ex4.png} &
		\includegraphics[height=5cm]{chat3.png} &
		\includegraphics[height=5cm]{chat4.png} \\
		(d) Our result & (e) GPT Mermaid & (f) GPT D2 \\
	\end{tabular}

	\caption{Six vertically-oriented images arranged in two rows}
	\label{fig:chat}

\end{figure}

Another result we'd like to present is a comparison between D.I.A.G.R.A.M.'s results and what GPT-4o could give in zero-shot. The prompt given was the following: "The current image features a diagram. Analyze it and output the respective D2 and Mermaid script.". The results given (see Figure~\ref{fig:chat}) may be - at first glance - quite similar. However, some key differences arise. For instance, in diagrams (b), (e), (f), GPT isn't able to correctly output the shape of the nodes - whereas our net is capable of doing. Just like our net added a node instead of the self-arrow (see preceding paragraph), GPT isn't able to correctly recognize that portion of ink as a self-arrow (images (e), (f)). Another subtle difference - this time, not logical - can be seen in image (c): GPT created three different nodes, while the original image had only one.
